---
title: |
  | An application of item response theory for agricultural sustainability measurement
  |
  | Estimation results
output:
  html_document:
    fig_caption: yes
    theme: spacelab
    highlight: pygments
    toc: yes
    toc_depth: 4
    number_sections: yes
    toc_float:
      smooth_scroll: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, warning=FALSE)
rm(list=ls())
library(tidyverse)
library(haven)
library(brms)
library(bayesplot)
library(ggplot2)
theme_set(theme_bw())
```

# Data

The Farm Accountancy Data Network (FADN) data used in this study are available officially only to registered users. For a full documentation and the conditions of access, see https://agriculture.ec.europa.eu/data-and-analysis/farm-structures-and-economics/fadn_en

Given access to the FADN original files, the data set can be
replicated using the supplementary information provided in this repository. 

```{r}
d <- read_dta("fadn-2013.dta")
d <- rename(d, id = idn)
d$id <- as.integer(d$id)
d$item <- as.factor(d$item)
table(d$item)
```


```{r}
d$y <- factor( d$y, levels = c("0", "1", "2", "3"), ordered = TRUE ) 
glimpse(d, width = 60)
table(d$y)
```

```{r}
summary(d)
```

# Baseline IRT model 

## Fitting 

```{r}
prior <-
  # discrimination
  prior("normal(0, 2)", class = "Intercept", dpar = "disc") + 
  prior("normal(0, 2)", class = "sd", group = "item", dpar = "disc") + 
  # unit-specific effects 
  prior("constant(1)", class = "sd", group = "id") +
  # item-specific effects 
  prior("normal(0, 2)", class = "sd", coef = "Intercept", group = "item") 
formula <- brmsformula(
  y ~ 1 +  (1|i|item) + (1|id),
  disc ~ 1 + (1|i|item)
)
fit1 <- brm( 
  formula = formula,
  data = d,
  family = cumulative(link = "logit"),
  prior = prior,
  cores = 4,
  iter = 1000,
  warmup = 500,
  refresh = 0,
  file = "results/01-baseline-IRT.rds"
)
```

```{r}
summary(fit1)
```

```{r}
prior_summary(fit1)
```

## Easiness and discrimination parameters (Figure 1)

This code replicates Figure 1 in the paper.

```{r}
# extract item and easiness parameters
ranef1 <- ranef(fit1)
# discriminations
alpha <- ranef1$item[, , "disc_Intercept"] %>%
  exp() %>% 
	as_tibble() %>%
	rownames_to_column()
# item easinesses (deviations from thresholds)
beta <- ranef1$item[, , "Intercept"] %>%
	as_tibble() %>%
	rownames_to_column()
# put easinesses and discriminations together
bind_rows(beta, alpha, .id = "nlpar") %>%
	rename(item = "rowname") %>%
	mutate(item = as.numeric(item)) %>%
	mutate(nlpar = factor(nlpar, labels = c("Easiness", "Discrimination"))) %>%
	ggplot(aes(item, Estimate, ymin = Q2.5, ymax = Q97.5)) +
	facet_wrap("nlpar", scales = "free_x") +
	geom_pointrange() +
	coord_flip() +
	labs(x = "Item Number") + 
  scale_x_discrete( limits=1:9 )
```

## Extract sustainability index (farm parameters)

```{r}
df.re <- ranef(fit1) 
dim(df.re$id) # dimension: no of units X no sum. stats X no. of parameters
df.re <- data.frame( asi = df.re$id[,1,] ) # extract mean
dim(df.re)
# convert row names into first column
df.re <- tibble::rownames_to_column(df.re, "id")
head(df.re)
```


Save sustainability indicator:

```{r}
nrow(df.re) # number of AS values: one value for each farm in each year 
# write.csv(df.re, file = "asi-index.csv")
```


# Modeling covariates: farm type classification (TF8)

## Fitting 

Create factor variables for use in the regression:

```{r}
d$ftype <- ifelse( d$TF8 == 1, 1,
                   ifelse( d$TF8 == 2, 2, 
                   ifelse( d$TF8 == 3, 3, 
                   ifelse( d$TF8 == 4, 4, 
                   ifelse( d$TF8 == 5, 5, 
                   ifelse( d$TF8 == 6, 6, 
                   ifelse( d$TF8 == 7, 7, 
                   ifelse( d$TF8 == 8, 8, NA))))))))
d$ftype <- factor(d$ftype,
                  levels = c(1,2,3,4,5,6,7,8),
                  labels = c("Fieldcrops", "Horticulture", "Wine", "Other permanent crops", 
                             "Milk", "Other grazing livestock", "Granivores", "Mixed"))
```

```{r}
prior <-
  # discrimination
  prior("normal(0, 2)", class = "Intercept", dpar = "disc") + 
  prior("normal(0, 2)", class = "sd", group = "item", dpar = "disc") +
  # unit-specific effects 
  prior("constant(1)", class = "sd", group = "id") +
  # item-specific effects 
  prior("normal(0, 2)", class = "sd", coef = "Intercept", group = "item") +
  # covariate effects 
  prior("normal(0, 2)", class = "b") 
formula <- brmsformula(
  y ~ 1 + ftype + (1|i|item) + (1|id),
  disc ~ 1 + (1|i|item)
)
fit1 <- brm( 
  formula = formula,
  data = d,
  family = cumulative(link = "logit"),
  prior = prior,
  cores = 4,
  iter = 1000,
  warmup = 500,
  refresh = 0,
  file = "results/02-farm-type.rds"
)
```

```{r}
summary(fit1)
```

```{r}
prior_summary(fit1)
```

## Predicted probabilities (Figure 2)

```{r}
foo <- conditional_effects( fit1, categorical = TRUE, prob = 0.8 )
plot(foo, plot = FALSE)[[1]] + 
  scale_x_discrete(name ="") +
  coord_flip() +
  theme_bw()
```

# Modeling covariates: economic class size

## Fitting

```{r}
prior <-
  # discrimination
  prior("normal(0, 2)", class = "Intercept", dpar = "disc") + 
  prior("normal(0, 2)", class = "sd", group = "item", dpar = "disc") +
  # unit-specific effects 
  prior("constant(1)", class = "sd", group = "id") +
  # item-specific effects 
  prior("normal(0, 2)", class = "sd", coef = "Intercept", group = "item") +
  # covariate effects 
  prior("normal(0, 2)", class = "b") 
formula <- brmsformula(
  y ~ 1 + size + (1|i|item) + (1|id),
  disc ~ 1 + (1|i|item)
)
fit1 <- brm( 
  formula = formula,
  data = d,
  family = cumulative(link = "logit"),
  prior = prior,
  cores = 4,
  iter = 1000,
  warmup = 500,
  refresh = 100,
  file = "results/03-size.rds"
)
```

```{r}
summary(fit1)
```

## Predicted probabilities (Figure 3)

```{r}
foo <- conditional_effects( fit1, categorical = TRUE, prob = 0.8 )
plot(foo, plot = FALSE)[[1]] + 
  scale_x_discrete(name ="") +
  coord_flip() +
  theme_bw()
```

# Modeling covariates: regions

## Fitting

```{r}
prior <-
  # discrimination
  prior("normal(0, 2)", class = "Intercept", dpar = "disc") +
  prior("normal(0, 2)", class = "sd", group = "item", dpar = "disc") +
  # unit-specific effects 
  prior("constant(1)", class = "sd", group = "id") +
  # item-specific effects 
  prior("normal(0, 2)", class = "sd", coef = "Intercept", group = "item") +
  # covariate effects 
  prior("normal(0, 2)", class = "b") 
formula <- brmsformula(
  y ~ 1 + NUTS2 + (1|i|item) + (1|id),
  disc ~ 1 + (1|i|item)
)
fit1 <- brm( 
  formula = formula,
  data = d,
  family = cumulative(link = "logit"),
  prior = prior,
  cores = 4,
  iter = 1000,
  warmup = 500,
  refresh = 0,
  file = "results/04-region.rds"
)
```

```{r}
summary(fit1)
```

```{r}
prior_summary(fit1)
```

## Predicted probabilities (Figure 4)

```{r}
newx <- data.frame( NUTS2 = unique(d$NUTS2) )
pp <- fitted(fit1, newdata = newx, re_formula = NA, probs = c(0.10, 0.90))
pp4 <- pp[,,"P(Y = very sustainable)"]
print(pp4)
```


```{r}
knitr::knit_exit()
```
